Lesson 5
========================================================
importing the data
```{r}
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
```

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes: She tried to see if there was a corelation between age and guesses, but seperating the box plot by age color didnt work due to scatteredness and overplotting

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
library(ggplot2)
library(dplyr)

#box plot
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + 
  geom_boxplot() +
  #adding mean to the box, x - mean
  stat_summary(fun.y = mean, shape = 4, geom = 'point')


#line
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) + 
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)


pf.fc_by_age_gender <- pf %>%
  group_by(age, gender) %>%
  filter(!is.na(gender)) %>%
  summarise(
     mean_friend_count = mean(friend_count),
     median_friend_count = median(friend_count),
     n = n()) %>%
  arrange(age)

```
 
***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y = median_friend_count), data = pf.fc_by_age_gender) +
  geom_line(aes(color = gender))
```

***

### Thinking in Ratios
Notes: Lets find the difference between friend count and gender by ratios ie: how many more times does x vs y

***

### Wide and Long Format
Notes: it might be helpful to convert the data into a wide format vs long

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)

pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender,
                                  value.var = 'median_friend_count')

pf.fc_by_age_gender.wide$fratio <- pf.fc_by_age_gender.wide$female / pf.fc_by_age_gender.wide$male

pf.fc_by_age_gender.wide$mratio <- pf.fc_by_age_gender.wide$male / pf.fc_by_age_gender.wide$female

head(pf.fc_by_age_gender.wide)




```
***

### Ratio Plot
Notes: I've created a fratio and mratio to see the difference in the ratio between gender. geom_hline yincercept is where you want there to be a line, alpha is the transparancy

```{r Ratio Plot}
ggplot(aes(x=age, y=fratio), data = pf.fc_by_age_gender.wide) +
  geom_line() +
  geom_hline(yintercept = 2, alpha = 1/1, linetype = 4)
```

***

### Third Quantitative Variable
Notes:
Create a variable called year_joined
in the pf data frame using the variable
tenure and 2014 as the reference year.

```{r Third Quantitative Variable}
library(date)
 
pf$year_joined <-  floor(2014 - (pf$tenure/365)) 

```
Notes:  got it!. floor() function rounds the variable to the nearest whole number, but it cannot be greater then the inputted number (ie: 5.6 will be 5)

***

### Cut a Variable
Notes: we can use c() and seq() to cut the data

(2004, 2009]
(2009, 2011]
(2011, 2012]
(2012, 2014]

```{r Cut a Variable}
?cut

Z <- pf$year_joined
pf$year_joined.bucket <- (cut(Z, c(2004,2009, 2011, 2012, 2014)))
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)

```

***

### Plot the Grand Mean
Notes:
rewatch this grand mean stuff
```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) + 
  geom_line(stat = 'summary', fun.y = mean, linetype = 2)

```

***

### Friending Rate
Notes: Find the rate at which people are adding friends per day

```{r Friending Rate}
friend_per_day <- with(data = subset(pf, tenure>0),
     friend_count/tenure)

summary(friend_per_day)

### how to simplify the code ###

with(subset(pf, tenure >= 1), summary(friend_count/tenure))


```

***

### Friendships Initiated
Notes:

What is the median friend rate?
.2205

What is the maximum friend rate?
417

notice how the max is 417, while the third quartile is .56. it is clear the max is an outlier

```{r Friendships Initiated}
# Create a line graph of mean of friendships_initiated per day (of tenure)
# vs. tenure colored by year_joined.bucket.

 ggplot(aes(x=tenure, y=friendships_initiated/tenure), data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary' ,fun.y = mean)
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket))

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv('yogurt.csv')
yo$id <- factor(yo$id)


ggplot(aes(x = price), data = yo) +
  geom_histogram()
```
what do you notice?
There are a lot of gaps between the bars. this is because the prices are set.  Most people buy @ the price of 65. 


***

### Number of Purchases
Notes: add all.purchases to the data using transform

```{r Number of Purchases}

?transform

yo <- transform(yo, all.purchases = strawberry + plain + blueberry + pina.colada + mixed.berry)

ggplot(aes(x = all.purchases), data = yo) +
  geom_histogram(binwidth = 1) 
```

***

### Prices over Time

```{r Prices over Time}
ggplot(aes(x = price, y = time), data = yo) +
  geom_point(alpha = 1/20)
```
Notes:

***

### Sampling Observations
Notes:  What i noticed is there are major purchases for a certain price. It could indicate a sale? 

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
?set.seed()
# set.seed() creates "fake" random numbers. it's good to use to simulate something. when trying to figer out a pattern (of "random" numbers) you what to reproduce, set.seed() creates a random number, but not random enough where you cannot recreate it.

?sample()
#takes a sample of the data

?levels()
# when the data has a level, this sets the levels into intergers (ie: a column can only contain  white, black, blue. that is a level. in this case the levels are the unique house holds.)

set.seed(200)
sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(y = price, x = time), data = subset(yo, id %in% sample.ids)) +
  geom_point(aes(size = all.purchases), shape = 1) +
  facet_wrap(~ id) +
  geom_line()
```
What did you ntoice? I noticed that most purchases were made at the "average" price.  When the purchase price was lower, I was surprised to see that more households did not buy more yogurts in bulk.

This is a good example of how you would plot something when you have data over time. We can see in the plots when some houseold bought a lot and a little.
***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes: a scatterplot matrix could be a good starting point because it graphs EVERYTHING and it can give us a quick overview of how the datas relate to each other.
```{r}
install.packages('GGally')
library(GGally)

theme_set(theme_minimal(20))
set.seed(1836)
pf_subset <- pf[c(2:15)]

names(pf_subset)
```

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

