---
title: "Life"
output: html_document
---
# The Gapminder website contains over 500 data sets with information about
# the world's population. Your task is to download a data set of your choice
# and create 2-5 plots that make use of the techniques from Lesson 3.

# You might use a simple histogram, a boxplot split over a categorical variable,
# or a frequency polygon. The choice is yours!

# You can find a link to the Gapminder website in the Instructor Notes.

# Once you've completed your investigation, create a post in the discussions that includes:
#       1. any questions you answered, your observations, and summary statistics
#       2. snippets of code that created the plots
#       3. links to the images of your plots

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Copy and paste all of the code that you used for
# your investigation, and submit it when you are ready.
# ====================================================================================

importing data and needed libraries
```{r}
library(rJava)
library(xlsxjars)
library(xlsx)
library(ggplot2)
library(plyr)
library(gridExtra)
library(grid)

library(tidyr)
library(dplyr)

life <- read.xlsx('life.xlsx', sheetIndex = 1)


# keep this just in case
names(life)[names(life)=="Life.expectancy"] <- "Countries"

life <- gather(life, 'year', 'age', 2:ncol(life), convert=T) #bring years down into rows with 'year' and 'age' columns

life$year <- factor(life$year, ordered=T) #Convert year to ordered factor


```
My tests
```{r}
myvars <- c("United States", "Japan", "United Kingdom")

test <-subset(life, Countries == 'United States')

test <- subset(life, Countries == 'United States' | Countries == 'Japan' | Countries == 'United Kingdom')

test <- t(test)
names(my_data)[names(my_data)=="Countries"] <- "Years"

my_data <- my_data[-1,]

colnames(test)[1] <- 'Japan'
colnames(test)[2] <- 'United_Kingdom'
colnames(test)[3] <- 'United_States'

head <- head(life)
my_data <- as.data.frame.matrix(my_data)
qplot(data = head, x= United_Kingdom)  

#geom_bar(stat = "identity")

```

Histograms
```{r}
# creates 3 plots, and saves them into the variables
x2016 <- qplot(data = life, x = X2016, bins = 50)
x1900 <- qplot(data = life, x = X1900, bins = 50)
x1800 <- qplot(data = life, x = X1800, bins = 50) 
```

# prints all 3 grid and assigns it to histo, saves it 
histo <- grid.arrange(x2016, x1900, x1800, ncol=2, top = 'Life Expectancy by year')
ggsave('life_histo.png', histo)


Box Plot
```{r}
qplot(data = life,x = X2000) +
  geom_freqpoly()
```

