---
title: "murder"
output: html_document
---

# The Gapminder website contains over 500 data sets with information about
# the world's population. Your task is to download a data set of your choice
# and create 2-5 plots that make use of the techniques from Lesson 3.

# You might use a simple histogram, a boxplot split over a categorical variable,
# or a frequency polygon. The choice is yours!

# You can find a link to the Gapminder website in the Instructor Notes.

# Once you've completed your investigation, create a post in the discussions that includes:
#       1. any questions you answered, your observations, and summary statistics
#       2. snippets of code that created the plots
#       3. links to the images of your plots

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Copy and paste all of the code that you used for
# your investigation, and submit it when you are ready.
# ====================================================================================


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing the data
```{r}
install.packages("xlsx", dep = T)  
install.packages('rJava')
install.packages('xlsxjars')
install.packages('plyr')
install.packages('reshape2')
library(rJava)
library(xlsxjars)
library(xlsx)
library(ggplot2)
library(plyr)
library(gridExtra)
library(grid)
library(reshape2)

#Imports the xlsx data into murders
murders <- read.xlsx('Murder.xlsx', sheetIndex = 1)

# renames the long col name to Countries
names(murders)[names(murders)=="Total.number.of.deaths.from.interpersonal.violence"] <- "Countries" # renames the long col name to Countries

#restructures murders into year and deaths
my_data <- gather(murders, 'year', 'deaths', 2:ncol(murders), convert=T) #bring years down into rows with 'year' and 'age' columns


#both murders and my_data are used for this lesson
``` 

Summaries
```{r}
summary(murders$X2002)
summary(murders$X2004)

summary(my_data$deaths)
```

Histograms
```{r}
y2004 <- qplot(x = log10(X2004), bins = 50, data = subset(murders, !is.na(X2004))) 
y2002 <- qplot(x = log10(X2002), bins = 50, data = subset(murders, !is.na(X2002))) 
total <- qplot(x = log10(deaths),xlab="total murders", bins = 50, data = subset(my_data, !is.na(murders)))

q<-grid.arrange(y2002, y2004, total, ncol=3)
ggsave('murder_histogram.png', q)
```

Box Plots
```{r}
box <-qplot(data = my_data, x = year, y = deaths, ylab = "Murder totals of all Countries") +
  geom_boxplot() +
  coord_cartesian(ylim = c(0,500))

box

ggsave("muder_box.png", box)
```

Freqency Chart
```{r}



top_10_2002 <- head(murders[order(-murders$X2002),], 10)
top_10_2004 <- head(murders[order(-murders$X2004),], 10)

top2002 <- gather(top_10_2002, year, deaths,2:ncol(top_10_2002), convert=T)

qplot(data = top2002, y = year, x= deaths, group=1) +
  geom_line(color = 'red')

two_years <- top_10_2002


line_2002 <- qplot(data = top_10_2002, x = X2002, y= Countries, group=1) +
  geom_line(color = 'red')

line_2004 <- qplot(data = top_10_2002, x = X2004, y= Countries, group=1) +
  geom_line(color = 'green')

line_2002
grid.arrange(line_2002, line_2004, ncol=1)


ggplot(data=top2002, aes(x=year, y=deaths, group = Countries, color = Countries)) +
    geom_line() 
ggsave('top10line.png')
```

