---
title: "Problem_set2"
output: html_document
---

Importing the data
```{r}
library(ggplot2)
data("diamonds")

```

Problem 1

# [x] Create a histogram of diamond prices.
# [x] Facet the histogram by diamond color
# and use cut to color the histogram bars.

# The plot should look something like this.
# http://i.imgur.com/b5xyrOu.jpg

# Note: In the link, a color palette of type
# 'qual' was used to color the histogram using
# scale_fill_brewer(type = 'qual')

# This assignment is not graded and
# will be marked as correct when you submit.

```{r}
ggplot(aes(x = log(price)), data = diamonds) +
  geom_histogram( aes(fill = cut)) +
  facet_wrap(~ color) 
#notice how we are using fill and not color in geom histogram
  
```


Problem 2

# Create a scatterplot of diamond price vs.
# table and color the points by the cut of
# the diamond.

# The plot should look something like this.
# http://i.imgur.com/rQF9jQr.jpg

# Note: In the link, a color palette of type
# 'qual' was used to color the scatterplot using
# scale_color_brewer(type = 'qual')

```{r}
ggplot(aes(x =  price, y = table), data = diamonds) +
  geom_point(aes(color = cut)) +
  scale_color_brewer(type = 'qual') 
```


Problem 4
# Create a scatterplot of diamond price vs.
# volume (x * y * z) and color the points by
# the clarity of diamonds. Use scale on the y-axis
# to take the log10 of price. You should also
# omit the top 1% of diamond volumes from the plot.

# Note: Volume is a very rough approximation of
# a diamond's actual volume.

# The plot should look something like this.
# http://i.imgur.com/excUpea.jpg

# Note: In the link, a color palette of type
# 'div' was used to color the scatterplot using
# scale_color_brewer(type = 'div')

# This assignment is not graded and
# will be marked as correct when you submit.

# ENTER YOUR CODE BELOW THIS LINE
# ===========================================
```{r}
ggplot(aes(y = price, x =(x*y*z)), data = diamonds) +
  geom_point(aes(color = clarity)) +
  coord_cartesian(xlim = c(0, quantile((diamonds$x*diamonds$y*diamonds$z), 0.99)))
  scale_y_log10()


?log10
```

Problem 5
# Many interesting variables are derived from two or more others.
# For example, we might wonder how much of a person's network on
# a service like Facebook the user actively initiated. Two users
# with the same degree (or number of friends) might be very
# different if one initiated most of those connections on the
# service, while the other initiated very few. So it could be
# useful to consider this proportion of existing friendships that
# the user initiated. This might be a good predictor of how active
# a user is compared with their peers, or other traits, such as
# personality (i.e., is this person an extrovert?).

# Your task is to create a new variable called 'prop_initiated'
# in the Pseudo-Facebook data set. The variable should contain
# the proportion of friendships that the user initiated.

# This programming assignment WILL BE automatically graded.

```{r}
pf <- read.csv("pseudo_facebook.tsv", sep = '\t')

pf$prop_initiated <- (pf$friendships_initiated/pf$friend_count)
pf$prop_initiated <- ifelse(is.na(pf$prop_initiated), 0, pf$prop_initiated)

ifelse(pf$prop_initiated)

pf$friend_count-pf$friendships_initiated/pf$friend_count
```

Problem 6
# Create a line graph of the median proportion of
# friendships initiated ('prop_initiated') vs.
# tenure and color the line segment by
# year_joined.bucket.

# Recall, we created year_joined.bucket in Lesson 5
# by first creating year_joined from the variable tenure.
# Then, we used the cut function on year_joined to create
# four bins or cohorts of users.

# (2004, 2009]
# (2009, 2011]
# (2011, 2012]
# (2012, 2014]

# The plot should look something like this.
# http://i.imgur.com/vNjPtDh.jpg
# OR this
# http://i.imgur.com/IBN1ufQ.jpg

# This assignment is not graded and
# will be marked as correct when you submit.

# ENTER YOUR CODE BELOW THIS LINE
```{r}
pf$year_joined <-  floor(2014 - (pf$tenure/365)) 
pf$year_joined.bucket <-  cut(pf$year_joined, c(2004, 2009, 2011, 2012, 2014))

ggplot(aes(x = tenure, y = prop_initiated), data = pf) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)
```

Problem 7

# Smooth the last plot you created of
# of prop_initiated vs tenure colored by
# year_joined.bucket. You can bin together ranges
# of tenure or add a smoother to the plot.

# There won't be a solution image for this exercise.
# You will answer some questions about your plot in
# the next two exercises.

# This assignment is not graded and
# will be marked as correct when you submit.

# ENTER YOUR CODE BELOW THIS LINE
# ====================================================
```{r}
ggplot(aes(x = tenure, y = prop_initiated), data = pf) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) +
  geom_smooth()

with(subset(pf, pf$year_joined > 2012), mean(prop_initiated))
```


problem 10

# Create a scatter plot of the price/carat ratio
# of diamonds. The variable x should be
# assigned to cut. The points should be colored
# by diamond color, and the plot should be
# faceted by clarity.

# The plot should look something like this.
# http://i.imgur.com/YzbWkHT.jpg.

# Note: In the link, a color palette of type
# 'div' was used to color the histogram using
# scale_color_brewer(type = 'div')

# This assignment is not graded and
# will be marked as correct when you submit.

# ENTER YOUR CODE BELOW THIS LINE
# ===========================================
```{r}
ggplot(aes(x = cut, y = price/carat), data = diamonds) +
  geom_jitter(aes(color = color)) +
  facet_wrap(~ clarity) +
   scale_color_brewer(type = 'div')

?geom_jitter
```

Try it!

```{r}
titanic <- read.csv('titanic_data.csv')

# average price paid per survivor, and their age, colored by class a

ggplot(subset(titanic, Survived == 1), aes(y = Fare, x = Age)) +
  geom_line(aes(color = Pclass), stat = 'summary', fun.y = mean) +
  facet_wrap(~Sex) +
  labs(title = 'Survivors')
       
ggplot(subset(titanic, Survived == 0), aes(y = Fare, x = Age)) +
  geom_line(aes(color = Pclass), stat = 'summary', fun.y = mean) +
  labs(title = 'Non-Survivors') +
  facet_wrap(~Sex)



names(titanic)
new_titanic <- titanic[,c(2:12)]


?ggpairs()
ggpairs(new_titanic)
```

