---
title: "Problem Set"
output: html_document
---

Imporing Data
```{r}
library(ggplot2)
library(dplyr)
data("diamonds")
?diamonds
```

Scatter plot
```{r}
ggplot(data = diamonds, aes(x=carat, y=price)) +
  geom_point(alpha = 1/20, position = position_jitter())
```
Observation: A majority of the 1.5 carat diamonds and under, are below 10,000 dollars in price. The data also forms vertical price lines indicating that there are a wide range of price per carat.

Correlation 
```{r}
# price and x
cor.test(diamonds$price, diamonds$x)

# price and y
cor.test(diamonds$price, diamonds$y)

# price and z
cor.test(diamonds$price, diamonds$z)
```

price and depth scatter plot
```{r}
ggplot(data = diamonds, aes(x = depth, y = price)) +
  geom_point(alpha = 1/20)
```


scatter plot of depth and price
```{r}
ggplot(data = diamonds, aes(x = depth, y = price)) + 
  geom_point(alpha = 1/100) +
  scale_x_continuous(breaks = seq(0,80,2))

```

correlation of depth and price
```{r}
cor.test(diamonds$depth, diamonds$price)
```

price and carot minus top 1%
```{r}

#two ways of doing it, one with subset
ggplot(data = subset(diamonds, price < quantile(price, c(.99))), aes(x=carat, y=price)) + geom_point(alpha = 1/100)


#one using limits
ggplot(data = diamonds, aes(x=carat, y=price)) + 
  geom_point(alpha = 1/100) +
  ylim(0,quantile(diamonds$price, .99))

```


Adding new column, volume!
```{r}
diamonds$volume = diamonds$x * diamonds$y * diamonds$z

ggplot(aes(x = volume, y = price), data = diamonds) +
  geom_point()
```
What do you observe? A great majority of the diamonds have a volume under 500 and a price under 10,000. We can see that there are some outliers. There are a few diamonds with 0 as their volume and one with a volume over 3,000.

Corelation of price and volume minus those with 0 volume
```{r}


with(diamonds[diamonds$volume != 0 & diamonds$volume < 800,], cor.test(price, volume))

#the line below was used to audit the data
r <- diamonds[diamonds$volume != 0 & diamonds$volume < 800,]

```


Plot no >=800 and no 0 volumes
```{r}
ggplot(aes(x = volume, y = price), data = subset(diamonds, volume !=0 & volume <800)) + geom_point(alpha = 1/100) +
  stat_smooth(method =  'lm') +
  xlim(0,800) +
  ylim(0,20000)

```

grouping by clarity and making a new dataframe
```{r}

diamondsByClarity <- diamonds %>%
  group_by(clarity)  %>%
  summarise( 
          mean_price = mean(price),
          median_price = median(price),
          min_price = min(price),
          max_price = max(price),
          n=n())  %>%
  arrange(clarity)
```

bar chars of mean prices
Create two bar plots and print it out in one image
```{r}
library(gridExtra)
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

p1 <- ggplot(aes(x = clarity, y = mean_price), data = diamonds_mp_by_clarity) +
  geom_bar(stat = 'identity')

p2 <- ggplot(aes(x = color, y = mean_price), data = diamonds_mp_by_color) +
  geom_bar(stat = 'identity')

diamonds_by_cut <- group_by(diamonds, cut)
diamonds_mp_by_cut <- summarise(diamonds_by_cut, mean_price = mean(price))

p3 <- ggplot(aes(x = cut, y = mean_price), data = diamonds_mp_by_cut) +
  geom_bar(stat = 'identity')

grid.arrange(p1,p2, p3, ncol=1)
```

Revisiting Gapminder
```{r}
library(rJava)
library(xlsxjars)
library(xlsx)
library(plyr)
library(reshape2)
library(ggplot2)
library(gridExtra)
library(grid)
library(tidyr)

murder <- read.xlsx('murder.xlsx', sheetIndex = 1)

# renames the long col name to Countries
names(murder)[names(murder)=="Total.number.of.deaths.from.interpersonal.violence"] <- "Countries" # renames the long col name to Countries

#restructures murders into year and deaths
my_data <- gather(murder, 'year', 'deaths', 2:ncol(murder), convert=T) #bring years down into rows with 'year' and 'age' columns

# replaces all NA value to 0
my_data[is.na(my_data)] <- 0

#makes new data with the following parameters 
new_data <- my_data %>%
  group_by(year)  %>%
  summarise( 
          mean_deaths = mean(deaths),
          median_death = median(deaths),
          min_death = min(deaths),
          max_deaths = max(deaths),
          n=n())

#bar graph
ggplot(aes(x=year, y=mean_deaths), data = new_data) +
  geom_bar(stat = 'identity')

ggplot(aes(x=year, y=deaths), data = my_data) +
  geom_point(alpha = 1/10)

```






